---
title: "p8105_hw2_tk2886"
author: "Tanvir Khan"
date: 2021-10-02
output: github_document
---

```{r setup, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```


# Problem 1 
### Importing Mr.Trash Wheel Data. Mr.Trash Wheel is an Excel File. 
```{r importing data with specific requirements}
trashwheel_data <- 
  read_excel(
    "./data/TrashWheel/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N408",
    )
```




### Cleaning Mr.Trash Wheel Data
```{r cleaning data}
trashwheel_data <-
    trashwheel_data %>%
    janitor::clean_names() 
```




### Using filter, mutate, as.integer to format and get specific data from Mr.Trash Wheel data
```{r getting specific data from Mr.Trash Wheel data by using filter, mutate, as.integer}
trashwheel_data <-
  trashwheel_data %>%
  filter(!str_detect(month, "Total")) %>%
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```


### Importing 2018 Precipitation Data. 2018 Precipitation is an Excel File. 
```{r importing 2018 Precipitation data with specific requirements}
precipitation18_data <-
  read_excel(
    "./data/TrashWheel/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B15")
```


### Cleaning 2018 Precipitation Data.
```{r cleaning 2018 precipitation data}
precipitation18_data <-
  precipitation18_data %>%
  janitor::clean_names() %>%
  mutate(year = "2018") %>%
  drop_na()
```



### Importing 2019 Precipitation Data. 2019 Precipitation is an Excel File, 
```{r importing 2019 Precipitation data with specific requirements}
precipitation19_data <-
  read_excel(
    "./data/TrashWheel/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B15")
```

### Cleaning 2019 Precipitation Data.
```{r cleaning 2019 precipitation data}
precipitation19_data <-
  precipitation19_data %>%
  janitor::clean_names() %>%
  mutate(year = "2019") %>%
  drop_na()
```




### Combining 2018 and 2019 Precipitation Data. Use bind_rows instead of r.bind. Never use r.bind. 
```{r combining 2018 and 2019 precipitation data}
precipitation_data <-
  bind_rows(precipitation18_data, precipitation19_data) %>%
  mutate(month = month.name[month]) %>%
  relocate(year, month, total) 
```




### Analysis of 2018 Precipitation Data and 2019 Precipitation Data
```{r storing important information for 2018 and 2019 precipitation }
variable_p18 = precipitation18_data %>% names()
variable_p19 = precipitation19_data %>% names()
total_precipitation18 = precipitation18_data %>% summarise(sum(total))
total_precipitation19 = precipitation19_data %>% summarise(sum(total))
median_p18 = precipitation18_data %>% summarise(median(total))
median_p19 = precipitation19_data %>% summarise(median(total))
mean_p18 = precipitation18_data %>% summarise(mean(total))
mean_p19 = precipitation19_data %>% summarise(mean(total))
p18_row = precipitation18_data %>% nrow()
p19_row = precipitation19_data %>% nrow()
p18_col = precipitation18_data %>% ncol()
p19_col = precipitation19_data %>% ncol()
```

After importing and cleaning the **2018 precipitation** data . The important variables names are **`r variable_p18`**. The total number of observation are **`r p18_row`**. There are important information that may be examined. The *total rows* in precipitation data in 2018 is **`r p18_row`**. The *total columns* in precipitation data in 2018 is **`r p18_col`**. The *total precipitation* in 2018 was **`r total_precipitation18`**. The *median precipitation* in 2018 was **`r median_p18`**. The *mean precipitation* in 2018 was **`r mean_p18`**.

After importing the **2019 precipitation** data. The important variables names are **`r variable_p19`**. The total number of observation are **`r p19_row`**. There are important information that may be examined. The *total rows* in precipitation data in 2019 is **`r p19_row`**. The *total columns* in precipitation data in 2018 is **`r p19_col`**. The *total precipitation* in 2019 was **`r total_precipitation19`**. The *median precipitation* in 2019 was **`r median_p19`**. The *mean precipitation* in 2019 was **`r mean_p19`**. 




### Analysis of Mr. Trash Wheel Data
```{r storing important information for Mr. Trash Wheel Data}
variable_trashwheel = trashwheel_data %>% names()
row_trashwheel = trashwheel_data %>% nrow()
col_trashwheel = trashwheel_data %>% ncol()
median_sportsball = trashwheel_data %>% 
  filter(year == 2019) %>% summarise(median(sports_balls))
```
After importing and cleaning the **Mr. Trash Wheel** data . The important variables names are **`r variable_trashwheel`**. The *total rows* in Mr.Trash Wheel data is **`r row_trashwheel`**. The *total columns* in Mr.Trash Wheel data is **`r col_trashwheel`**. The *median number of sports ball* is **`r median_sportsball`**.





# Problem 2
### Importing and cleaning the FiveThirtyEight Data. 
```{r message = FALSE, warning = FALSE}
pol_data <- 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year","month", "day")) %>%
  mutate(across(.cols = c(year, month, day), as.integer)) %>%
  mutate(month = month.name[month]) %>%
  mutate(president = case_when(prez_dem == 1 ~ "dem", 
                               prez_gop == 1 ~ "gop",
                               prez_gop == 2 ~ "gop")) %>%
  select(year, month, president, everything(), -day, -prez_dem, -prez_gop)
```

### Importing and cleaning the snp - Stock Market Data.
```{r message = FALSE, warning = FALSE}
stockmark_data <-
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, sep = "/", into = c("month", "day", "year")) %>%
  mutate(across(.cols = c(month, day, year), as.integer)) %>%
  mutate(month = month.name[month])  %>%
  mutate(year = ifelse(year > 21, 1900 + year, 2000 + year)) %>%
  select(year, month, close)
```

### Importing and cleaning the unemployment Data.
```{r message = FALSE, warning = FALSE}
unemploy_data <-
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percentage"
  ) %>%
  mutate(across(.col = c(year), as.integer)) %>%
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June",                                "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" =                                  "December"))
```


### Combining the "pols-month" data and "snp" Data by using left join. 
```{r comibing the pols-month data and snp-data}
merge_data1 <-
  left_join(pol_data, stockmark_data, by = c("year", "month"))
```


### Combining the newly constructed data (merged_data1) with the umeployment data by using left join. 
```{r combiing the newly constructed data with unemployment data}
merge_data2 <-
  left_join(merge_data1, unemploy_data, by = c("year", "month"))
```


### Analysis of FiveThirtyEight Datas 
```{r Analyzing important variables in the FiveThirtyEight Data by storing them for later use}
name_poldata = pol_data %>% names()
dim_poldata = pol_data %>% dim()
range_year = pol_data %>% 
  select(year) %>%
  range()

name_stockmarket = stockmark_data %>% names()
dim_stockmarket = stockmark_data %>% dim()
range_year_snp = stockmark_data %>% 
  select(year) %>%
  range()

names_unemployment = unemploy_data %>% names()
dim_unemployment = unemploy_data %>% dim()
range_year_unemployment = unemploy_data %>% 
  select(year) %>%
  range()
```
The *"pols-month"* dataset contains data that have national politicians who are democratic or republican. The important variables in the pols-month dataset are `r name_poldata`. The dimensions of the pols-month data is `r dim_poldata`. The range of years in the pols-month data is from `r range_year`. 

The *“snp”* dataset contains data that is related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. The important variables in the snp dataset are `r name_stockmarket`. The dimensions of the snp data is `r dim_stockmarket`. The range of years in the snp data is from `r range_year_snp`. 

The *"unemployment"* data set contains data percentage of unemployment for a specific year and month. The important variables in the unemployment are `r names_unemployment`. The dimensions in the uneployment data set are `r dim_unemployment`. The range of years in the unemployment is from `r range_year_unemployment`. 


### Problem 3
### Importing and cleaning the Baby Name data. 
```{r message = FALSE, warning = FALSE}
babyname_data <-
  read_csv("./data/BabyNames/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(gender = str_to_sentence(gender),
         ethnicity = recode(ethnicity, "WHITE NON HISP" = "WHITE NON HISPANIC",
                           "BLACK NON HISP" = "BLACK NON HISPANIC",
                           "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                           "ASIAN AND PACIFIC" = "ASIAN AND PACIFIC ISLANDER"),
         ethnicity = str_to_title(ethnicity),
         childs_first_name = str_to_sentence(childs_first_name))
```

### Removing duplicate rows from Baby Name data.
```{r removing duplicate rows}
babyname_data <-
  babyname_data %>% distinct()
```

### Olivia Rank Data table 
```{r creating a table showing the rank popularity of the name -Olivia}
Olivia_rank <- 
  filter(babyname_data, childs_first_name == "Olivia" & gender == "Female") %>%
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>%
  select(childs_first_name, gender, ethnicity, "2011", "2012", "2013", "2014", "2015", "2016")

knitr::kable(Olivia_rank)
```

### Creating Male Rank Table 
```{r creating Male Baby Name rank table}
Male_rank <-
  filter(babyname_data, gender == "Male" & rank == 1) %>%
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>%
  select(rank, gender, ethnicity, "2011", "2012", "2013", "2014", "2015", "2016")

knitr::kable(Male_rank)
```

### Scatterplot Data 
```{r Extracting a specific data from the Popular name data set and this data will be later used to create a scatterplot}
scatterplot_data <-
  filter(babyname_data, gender == "Male", year_of_birth == "2016", ethnicity == "White Non Hispanic") 
```


### Creating Scatterplot
```{r Creating a scatterplot with the extracted data}
male_wnh16 = ggplot(scatterplot_data, aes(x = rank, y = count)) +
  geom_point() + 
  ggtitle("Male Non-White Hispanic Male Children Name in Popularity Rank Versus Real Count") + theme(plot.title = element_text(hjust = 0.5))

male_wnh16
```


### Saving my second ggplot (scatterplot) as PDF
```{r Saving ggplot as PDF, message = FALSE, warning = FALSE}
ggsave("HW2scatter_plot.pdf", plot = male_wnh16, height = 4, width = 6)
```
